VM-Centric Snapshot Deduplication for Converged Cloud Architectures

### 2014-07-19 snapshot_system/paper_dedup.tex
re

We describe our overall deduplication approach as consisting of three
complementary strategies: VM-local
duplicate search, global deduplication using popular chunks, and 
VM-centric file system block management. 
\begin{itemize}
\item 
\textbf{Cross-VM global deduplication using popular chunks} --
We separate the deduplication within a VM and cross VMs
and simplify the cross-VM deduplication while maximizing the inter-VM deduplication efficiency  as much as possible.
This is because global deduplication that detects the appearance of a chunk 
in any VM requires a substantial resource for fingerprint comparison.
To simplify cross-VM deduplication, we restrict the search scope of deduplication within the top $k$ 
most popular items. This popular data set is called the \textbf{PDS}. 
%This step accomplishes the canonical global fingerprint lookup using a popular fingerprint index.
 \begin{figure}
 \centering
  \includegraphics[width=3.25in]{figures/zipf_count_rank.pdf}
 \caption{Duplicate frequency versus  chunk ranking in a log scale after local deduplication.}
 \label{fig:Datazipf}
 \end{figure}

% {\it [Need to find a place to put these numbers in: Total number of chunks 
% in 350 snapshots: 1,546,635,485. 
% Total number of chunks after localized dedup: 283,121,924. Total number of unique chunks: 87,692,682.]}
The observation that motivates us to use the PDS
is that for those chunks that appear in different VMs, the top $k$ most popular items
dominate the appearance distribution.  
Figure~\ref{fig:Datazipf} shows the distribution of chunk popularity for a data trace 
from Alibaba with 2500 VMs discussed in Section ~\ref{sect:evaluation}.
We define chunk popularity as the number of unique copies of the chunk in the data-store,
i.e., the number of copies of the chunk after local deduplication.
%the distribution of chunk popularity~\cite{WeiZhangIEEE} 
%based on a production data trace from Alibaba with over 2500 VMs.
The distribution from this figure is Zipf-like. 
Let  $\sigma$ denote as the percentage of unique data chunks belonging to PDS and 
from the evaluation in Section~\ref{sect:evaluation}, we find that
$\sigma$ with about  2\% can deliver a fairly competitive deduplication efficiency.
%$\sigma$ with about  range of 2 to 4\% can deliver a fairly competitive deduplication efficiency.
%For example,
%with top 2\% of the most popular chunks stored in PDS, up-to 40\% of the raw backup 
%data can be eliminated by fingerprint lookup in PDS index.

@@

Since the previous work typically uses a non-uniform chunk size 
with an average of 4KB or 8KB for the best deduplication 
effectiveness~\cite{Guo2011,extreme_binning09,bottleneck08,Dong2011},
we conduct additional local similarity guided deduplication 
by comparing chunk fingerprints of a dirty segment on a snapshot 
with those in  {\em potentially similar} segments from its parent snapshot. 
We consider a parent  segment is  potentially similar to the current dirty segment if 1) the parent segment
is at the same offset as the current segment.
2) the signature of these segments is the same.
The signature of a segment is defined as the minimum value of all its chunk fingerprints 
computed during backup and is recorded in the snapshot metadata (called recipe). 
%Note that this definition of content similarity is an approximation~\cite{resemblance97}.  
This signature definition is motivated by the previous work for similarity-based routing in parallel
deduplication~\cite{extreme_binning09,Dong2011} and near duplicate detection~\cite{shingling97}.

When processing a dirty segment,
its  similar segments can be found easily from the
parent snapshot metadata.  Then metadata of the similar segments is loaded to memory,
which contain chunk fingerprints to be compared.
To control the time cost of search, we set a limit on the number of  similar segment recipes to be fetched. 
% Then,
%given a set of data chunks within a dirty segment,  we compare  these chunk fingerprints
%with those in similar segments.  
For example, assume that  a segment is of size  2MB, 
its segment recipe is roughly 19KB which contains about 500 chunk fingerprints and other chunk metadata.
By limiting at most 10 similar segments to search, the amount of memory for maintaining those 
similar segment recipes is 190K, which is insignificant. 
%compared to other memory requirements.

>>>>

We describe our overall deduplication approach as consisting of three
complementary strategies: VM-local
duplicate search, global deduplication using popular chunks, and 
VM-centric file system block management. 
\begin{itemize}
\item 
\textbf{Cross-VM global deduplication using popular chunks} --
We separate the deduplication within a VM and cross VMs
and simplify the cross-VM deduplication while maximizing the inter-VM deduplication efficiency  as much as possible.
This is because global deduplication that detects the appearance of a chunk 
in any VM requires a substantial resource for fingerprint comparison.
To simplify cross-VM deduplication, we restrict the search scope of deduplication within the top $k$ 
most popular items. This popular data set is called the \textbf{PDS}. 
%This step accomplishes the canonical global fingerprint lookup using a popular fingerprint index.
 \begin{figure}
 \centering
  \includegraphics[width=3in]{figures/zipf_count_rank.pdf}
 \caption{Duplicate frequency versus  chunk ranking in a log scale after local deduplication.}
 \label{fig:Datazipf}
 \end{figure}

% {\it [Need to find a place to put these numbers in: Total number of chunks 
% in 350 snapshots: 1,546,635,485. 
% Total number of chunks after localized dedup: 283,121,924. Total number of unique chunks: 87,692,682.]}
What  motivates us to use PDS
is that for those chunks that appear in different VMs, the top $k$ most popular items
dominate the appearance distribution.  
Figure~\ref{fig:Datazipf} shows the distribution of chunk popularity for a data trace 
from Alibaba with 2500 VMs discussed in Section ~\ref{sect:evaluation}.
We define chunk popularity as the number of unique copies of the chunk in the data-store,
i.e., the number of copies of the chunk after local deduplication.
%the distribution of chunk popularity~\cite{WeiZhangIEEE} 
%based on a production data trace from Alibaba with over 2500 VMs.
The distribution from this figure is Zipf-like. 
Let  $\sigma$ denote as the percentage of unique data chunks belonging to PDS and 
from the evaluation in Section~\ref{sect:evaluation}, we find that
$\sigma$ with about  2\% can deliver a fairly competitive deduplication efficiency.
%$\sigma$ with about  range of 2 to 4\% can deliver a fairly competitive deduplication efficiency.
%For example,
%with top 2\% of the most popular chunks stored in PDS, up-to 40\% of the raw backup 
%data can be eliminated by fingerprint lookup in PDS index.

@@

Since the previous work typically uses a non-uniform chunk size 
with an average of 4KB or 8KB for the best deduplication 
effectiveness~\cite{Guo2011,extreme_binning09,bottleneck08,Dong2011},
we conduct additional local similarity guided deduplication 
by comparing chunk fingerprints of a dirty segment in a snapshot 
with those in  {\em potentially similar} segments from its parent snapshot. 
We consider a parent  segment is  potentially similar to the current dirty segment if 1) the parent segment
is at the same offset as the current segment.
2) the signature of these segments is the same.
The signature of a segment is defined as the minimum value of all its chunk fingerprints 
computed during backup and is recorded in the snapshot metadata (called recipe). 
%Note that this definition of content similarity is an approximation~\cite{resemblance97}.  
This signature definition is motivated by the previous work for similarity-based routing in parallel
deduplication~\cite{extreme_binning09,Dong2011} and near duplicate detection~\cite{shingling97}.

When processing a dirty segment,
its  similar segments can be found easily from the
parent snapshot metadata.  Then metadata of the similar segments is loaded to memory,
containing chunk fingerprints to be compared.
To control the time cost of search, we set a limit on the number of  similar segment recipes to be fetched. 
% Then,
%given a set of data chunks within a dirty segment,  we compare  these chunk fingerprints
%with those in similar segments.  
For example, assume that  a segment is of size  2MB, 
its segment recipe is roughly 19KB which contains about 500 chunk fingerprints and other chunk metadata.
By limiting at most 10 similar segments to search, the amount of memory for maintaining those 
similar segment recipes is 190K, which is insignificant. 
%compared to other memory requirements.

### 2014-07-19 snapshot_system/paper_del.tex
re

\begin{figure}[htbp]
\vspace{2em}
  \centering
  \epsfig{file=images/deletion.pdf, width=3.35in}
  \caption{Fast approximate deletion}
  \label{fig:deletion_flow}
\end{figure}

Snapshot deletions can occur frequently since old snapshots become less useful.
General deduplication complicates the deletion process because sharing of duplicates
requires a global reference counting~\cite{Guo2011,Fabiano2013} 
 to identify if  a chunk can be  safely removed without any reference.
%Since our VM-centric design restricts sharing of data chunks only to a small dataset,
%we can greatly simplify the deletion process by 
%focusing on  unreferenced non-PDS chunks within each VM. 
%This process can be conducted independently per VM, and thus results in  a simpler flow control and 
%lower resource usage. 
One of the key benefits of our VM-centric design is that it allows for faster and simpler deletion. 
We propose the following strategies to speedup the deletion process and reduce its resource usage.
\begin{itemize}
\item First, since
sharing in our approach is limited to only a small dataset, we can greatly simplify deletion by focusing
on intra-VM sharing rather than inter-VM sharing. Then the snapshot
 deletion can therefore be conducted independently among VMs.
\item Second,  we propose an {\em approximate} deletion strategy to trade deletion accuracy for speed and
resource usage. Since the VM size is highly skewed in practice,  a large VM may still require
a substantial amount of memory  to facilitate the mark-sweep process of data chunks used within all snapshots
of this VM. 
Our approximate deletion method sacrifices a small percent of storage leakage
to efficiently identify unused chunks using a per-snapshot bloom-filter.
\end{itemize}
The simplified deletion algorithm contains three aspects.
\begin{itemize}
\item {\bf Computation for snapshot reference summary.}
Every time there is a new snapshot created,
we compute a Bloom-filter with $z$ bits as the reference summary vector for all non-PDS chunks used 
in this snapshot.
The items we put into the summary vector are all the references appearing in the metadata of the snapshot.
For each VM we preset the vector size according to estimated VM image size;
given $h$ snapshots stored for a VM, there are $h$ summary vectors maintained.
We adjust the summary vector size and recompute the vectors if the VM size changes substantially over time.
This can be done during periodic leakage repair described below.

@@

%{\bf Discussion}
We estimate storage leakage size and how often leak repair should be conducted.
Assume that  a VM keeps $h$ snapshots in backup storage, creates and deletes one snapshot
every day. Let $u$ be the number of chunks brought by initial backup for a VM, $\Delta u$ be the average
number of additional chunks added from one version to next snapshot version. Then the total number of 
chunks in a VM's snapshot store is about:
\[
U = u + (h-1)\Delta u.
\]

>>>>

\begin{figure}[htbp]
%\vspace{2em}
  \centering
  \epsfig{file=images/deletion.pdf, width=3.35in}
  \caption{Fast approximate deletion}
  \label{fig:deletion_flow}
\end{figure}

Snapshot deletions can occur frequently since old snapshots become less useful.
Both space accounting and reclamation are difficult because deduplication
greatly complicates how we identify unused data when deletion occurs.
To filter out un-referenced data from shared duplicates,
it would require to either maintain expensive live schemas or conduct
global reference counting~\cite{Guo2011}.
%or use some advanced mark-and-sweep methodologies
%which are all unsuitable for converged architecture as discussed in section~\ref{sect:background}.
% to identify if a chunk can be safely removed without any reference.
%Since our VM-centric design restricts sharing of data chunks only to a small dataset,
%we can greatly simplify the deletion process by 
%focusing on  unreferenced non-PDS chunks within each VM. 
%This process can be conducted independently per VM, and thus results in  a simpler flow control and 
%lower resource usage. 
One of the key benefits of our VM-centric design is that it allows for faster and simpler deletion. 
We propose the following strategies to speedup the deletion process and reduce its resource usage.
\begin{itemize}
\item First, since
sharing in our approach is limited to only a small dataset, we can greatly simplify deletion by focusing
on intra-VM sharing rather than inter-VM sharing. Then the snapshot
 deletion can therefore be conducted independently among VMs.
\item Second,  we propose an {\em approximate} deletion strategy to trade deletion accuracy for speed and
resource usage. Since the VM size is highly skewed in practice,  a large VM may still require
a substantial amount of memory  to facilitate the mark-sweep process of data chunks used within all snapshots
of this VM. 
Our approximate deletion method sacrifices a small percent of storage leakage
to efficiently identify unused chunks using a per-snapshot bloom-filter.
\end{itemize}
The simplified deletion algorithm contains three aspects.
\begin{itemize}
\item {\bf Computation for snapshot reference summary.}
Every time there is a new snapshot created,
we compute a Bloom-filter with $z$ bits as the reference summary vector for all non-PDS chunks used 
in this snapshot.
The items we put into the summary vector are all the references appearing in the metadata of the snapshot.
For each VM we preset the vector size according to estimated VM image size;
given $h$ snapshots stored for a VM, there are $h$ summary vectors maintained.
We adjust the summary vector size and recompute the vectors if the VM size changes substantially over time.
This can be done during periodic leakage repair described below.

@@

%{\bf Discussion}
We estimate storage leakage size and how often leak repair should be conducted.
Assume that  a VM keeps $h$ snapshots in backup storage, creates and deletes one snapshot
every day. Let $u$ be the number of chunks brought by initial backup for a VM, $\Delta u$ be the average
number of additional chunks added from one version to next snapshot version. Then the total number of 
chunks in a VM's snapshot store is about:
$
U = u + (h-1)\Delta u.
$
